% readme.tex -- a short example of how each Stata Journal insert should be
% organized.

\inserttype{article}

\author{Short article author list}{%
	Kerui Du\\
	School of Management\\
	Xiamen University\\
	Xiamen, China\\
	kerrydu@xmu.edu.cn \\
	\\
	
	\and
	Luis Orea \\
	Department of Economics \\
	School of Economics and Business\\
	University of Oviedo \\
	Oviedo, Spain \\
	lorea@uniovi.es \\
	\\
	
	\and
	Inmaculada C. Álvarez \\
	Department of Economics \\
	Universidad Autónoma de Madrid \\
	Madrid, Spain\\
	inmaculada.alvarez@uam.es
  
  
}

\title[Short toc article title]{Fitting spatial stochastic frontier models in Stata}
\maketitle


\begin{abstract}
In this article, we introduce a new command spxtsfa for fitting spatial stochastic frontier models in Stata. Over the last decades, an important theoretical progress of stochastic frontier models is the incorporation of various types of spatial components. Models with the ability to account for spatial dependence and spillovers have been developed for efficiency and productivity analysis, drawing extensive attention from industry and academia. Due to the unavailability of the statistical packages, the empirical applications of the new stochastic frontier models appear to be lagging. The spxtsfa command provides a routine for estimating the spatial stochastic frontier models in the style of \cite{orea2019new} and \cite{galli2022spatial}, enabling users to handle different sources of spatial dependence. In the presented article, we introduce the spatial stochastic frontier models, describe the syntax and options of the new command, and provide several examples to illustrate its usage.
	
	\keywords{stochastic frontier models, SFA, spatial dependence, technical efficiency, spillovers }
\end{abstract}


% discussion of the Stata Journal document class.
%\input sj.tex
% discussion of the Stata Press LaTeX package for Stata output.
%\input stata.tex
\input{section1-2.tex}
% discussion of the Stata Press LaTeX package for Stata output.
\input{section3-6.tex}


\bibliographystyle{sj}
\bibliography{sj}


\begin{aboutauthors}

Kerui Du is an associate professor at the School of Management, Xiamen University. His primary research interests include applied econometrics, energy and environmental economics.	

Luis Orea is a full professor at the School of Economics and Business, University of Oviedo. His primary research interests include Efficiency and productivity analysis, econometric modelling, agricultural economics, energy economics, regulation and competition, spatial economics. 

Inmaculada C. Álvarez is a full professor at the Department of Economics, Universidad Autónoma de Madrid. Her primary research interest include infrastructures, efficiency and productivity, economic growth and development, spatial economics and quantitative methods. 
	

	
\end{aboutauthors}


\endinput
